---
title: "Data"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
date: "2022-11-24"
---


```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(dplyr)
knitr::opts_chunk$set(
  fig.height = 6,
  fig.width = 8,
  message = F,
  echo = T,
  warning = F
)
```
## Data Sources

*  The second data set was downloaded from the [NHANES](https://www.cdc.gov/nchs/nhanes/index.htm) database.



## Data Processing and Cleaning

*  For data from the NHANES database:


```{r}
demo=foreign::read.xport("nhanes/P_DEMO.XPT") %>%
  janitor::clean_names() %>%
  select(seqn,riagendr,ridageyr,ridreth3,dmdeduc2,indfmpir) %>%
  drop_na() %>%
  rename(gender=riagendr,
         age=ridageyr,
         race=ridreth3,
         education_level=dmdeduc2,
         income_poverty_ratio=indfmpir) %>%
  mutate(gender=case_when(gender==1 ~ "male",
                          gender==2 ~ "female")) %>%
  mutate(race=case_when(race==1 ~ "Mexican American",
                        race==2 ~ "Other Hispanic",
                        race==3 ~ "Non-Hispanic White",
                        race==4 ~ "Non-Hispanic Black",
                        race==6 ~ "Non-Hispanic Asian",
                        race==7 ~ "Other Race")) %>%
  mutate(education_level=case_when(education_level==1 ~ "less than 9th grade",
                                   education_level==2 ~ "9-11th grade",
                                   education_level==3 ~ "high school graduate",
                                   education_level==4 ~ "college or AA degree",
                                   education_level==5 ~ "college graduate or above",
                                   education_level==7 ~ "refused",
                                   education_level==9 ~ "don't know"))


sleep_df=foreign::read.xport("nhanes/P_SLQ.XPT") %>%
  janitor::clean_names() %>% 
  select(seqn,slq300,slq310,sld012,slq320,slq330,sld013,slq050,slq120) %>%
  rename(weekday_slp_time=slq300,
         weekday_wake_time=slq310,
         weekday_slp_hr=sld012,
         weekend_slp_time=slq320,
         weekend_wake_time=slq330,
         weekend_slp_hr=sld013,
         trouble_slp=slq050,
         sleepy_freq=slq120
         )              %>%
  filter(weekday_slp_time!=77777 & weekday_slp_time!= 99999) %>%
  filter(weekday_wake_time!=77777 & weekday_wake_time!= 99999) %>%
  filter(weekend_slp_time!=77777 & weekend_slp_time != 99999) %>%
  filter(weekend_wake_time!= 77777 & weekend_wake_time != 99999) %>%
  filter(trouble_slp!=7 & trouble_slp!= 9) %>%
  filter(sleepy_freq!= 7 & sleepy_freq!=9) %>%
  drop_na() %>%
  mutate(trouble_slp=case_when(trouble_slp==1 ~ "yes",
                               trouble_slp==2 ~ "no")) %>%
  mutate(sleepy_freq=case_when(sleepy_freq==0 ~ "never",
                               sleepy_freq==1 ~ "once a month",
                               sleepy_freq==2 ~ "2-4 per month",
                               sleepy_freq==3 ~ "5-15 per month",
                               sleepy_freq==4 ~ "16-30 per month",
                               ))

```

```{r}
slp_df=inner_join(demo,sleep_df,by="seqn") %>%
  na_if("") %>%
  na.omit()
write_csv(slp_df, "data/slp_df.csv")
```


## Data Description



#### Social factors related to Sleeping Status Data

The resulting data file of `slp_df` contains a single dataframe with `r nrow(slp_df)` rows of data on `r ncol(slp_df)` variables, the list below is our variables of interest:



  * `seqn`. The sequence number of the candidate.
  * `gender`. Gender of the candidate.
  * `age`. Age of the candidate.
  * `race`. Race of the candidate.
  * `hour`. Time(hour) violation occurred.
  * `education_level`. The education level of the candidate.
  * `income_poverty_ratio`. A ratio to classify poverty and not poverty. Ratio greater than 1 will be considered as not        poverty, ratio less than 1 will considered as poverty.
  * `weekday_slp_time`. Sleep time on weekdays or workdays.
  * `weekday_wake_time`. Wake time on weekdays and workdays.
  * `weekday_slp_hr`. Average sleeping hours on weekdays or workdays.
  * `weekend_slp_time`. Sleep time on weekends.
  * `weekday_wake_time`.Wake time on weekends.
  * `weekend_slp_hr`. Average sleeping hours on weekends.
  * `trouble_slp`. A response on whether telling the doctor having trouble sleeping.
  * `sleepy_freq`. Average sleeply times during daytime per month.
  
