
---
title: "Final Project Proposal"
author: Jingchen Chai, Wenjia Zhu, Ruihan Zhang(rz2614), Yifei Zhao, Chenyao Ni(cn2604)
output: github_document
---

```{r include=FALSE}
library(tidyverse)
```

## **Factors Influencing Sleeping**
## Motivation of our project
Adults should sleep 7 or more hours per night on a regular basis to promote optimal health.
However, more than one-third of the US population gets less than the recommended amount of sleep. Sleep is a necessary physiological process for individuals, interacting with several economic, social and physical factors, influencing risks of multiple health problems. Therefore, based on NHANES dataset, we focus on sleep-related public health issues to carry out our research, hoping to popularize the significance of adequate and healthy sleep, arousing the public's health awareness of improving sleep quality and obtaining adequate sleep duration. We also want to visualize and discuss the biological, economic and social factors interacting with sleep, informing the public and healthcare providers on the importance of advocating public policy to promoting the nowadays sleep situation and overall public health.


## Intended Final Project
A report analyzing the prevalence of insufficient sleep in relation to socioeconomic and physiological factors, and an interactive panel to predict the risk of insufficient sleep for individuals. 

## Anticipated Data Resources

We plan to utilize data from the [CDC](https://www.cdc.gov/sleep/data-and-statistics/adults.html) website. The data are based on the Behavioral Risk Factor Surveillance System (BRFSS) survey. The data we are using are the age-adjusted prevalence of adults aged greater than 18 who reported short sleep duration varied by state in 2020. We plan to focus on the states, and create a map according to the distributions of the age-adjusted prevalence among different states. 

The second data set we will use is a [questionnaire sleeping data](https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/SLQ_J.htm) and also the date related with their [demographic](https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/DEMO_J.htm) information, both comes from the [NHANES](https://www.cdc.gov/nchs/nhanes/index.htm) website. We plan to merge the two data sets, and include variables: 

The third data set is [PLACES: Census Tract Data](https://catalog.data.gov/dataset/places-census-tract-data-gis-friendly-format-2020-release-fb1ec). 



## The planned analyses / visualizations / coding challenges
We plan to start from a map GIF image, which illustrates prevalence of short sleep duration among US adults by state from 2016 to 2020. 

For NHANES datasets, firstly, We can simply build a linear model between sleep disorder factors and length of sleep, and from this model we could briefly obtain the sleep length with sleee behavioral factors. Secondly, we classify the people into two groups by length of sleep in weekends and weekdays, people with sufficient sleep and people with insufficient sleep. We plan to use scatterplot and barplot to show the basic features of sleep conditions of the data. Thirdly, after joint with other datasets, we could discover the correlation among sleep conditions and social economic factors using bubble charts and heatmap.

As for PLACES: Census Tract and City Data, we firstly classify the data into two groups by different geographic level, census tract and city, and use the same method to analyze and compare the two groups in further study. For each group, we focus on the key variables like cityname, category, disease name, datavalue, geolocation and populationcount. Visulization method will include Leaflet for mapping with corrsponding category and populationcount, correlation plots among sleep datavalue and other disease datavalues, and linear model for such factors.

The coding challenges are mainly from tidying data and do further manipulation like aggregating and creating linear models, since some values are missing, undefined or ourliers, and the models maybe irrational if it is simply based on subjective factors.

## Timeline
```{r echo=FALSE}
timeline_df = read.csv("timeline.csv")
timeline_df %>% knitr::kable()
```

