---
title: "Sleep and Social Factors"
output: 
  html_document:
    toc: false
    toc_float: false
    code_folding: hide
          
---
On this page, we will be exploring the overall trends for our key outcome- sleeping hours per day, and how these outcomes correlate with the predictors, including demographic variables like age, race, sex, education level, and poverty status.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,
                      warning=FALSE, 
                      fig.width = 6,
                      fig.asp = .6,
                      out.width = "90%")
```

```{r}
library(tidyverse)
library(patchwork)
library(knitr)
library(dplyr)
library(gganimate)
library(gifski)
library(png)
library(plotly)
library(ggridges)

```


```{r,echo=FALSE}
demo=foreign::read.xport("nhanes/P_DEMO.XPT") %>%
  janitor::clean_names() %>%
  select(seqn,riagendr,ridageyr,ridreth3,dmdeduc2,indfmpir) %>%
  drop_na() %>%
  rename(gender=riagendr,
         age=ridageyr,
         race=ridreth3,
         education_level=dmdeduc2,
         income_poverty_ratio=indfmpir) %>%
  mutate(gender=case_when(gender==1 ~ "male",
                          gender==2 ~ "female")) %>%
  mutate(race=case_when(race==1 ~ "Mexican American",
                        race==2 ~ "Other Hispanic",
                        race==3 ~ "Non-Hispanic White",
                        race==4 ~ "Non-Hispanic Black",
                        race==6 ~ "Non-Hispanic Asian",
                        race==7 ~ "Other Race")) %>%
  mutate(education_level=case_when(education_level==1 ~ "less than 9th grade",
                                   education_level==2 ~ "9-11th grade",
                                   education_level==3 ~ "high school graduate",
                                   education_level==4 ~ "college or AA degree",
                                   education_level==5 ~ "college graduate or above",
                                   education_level==7 ~ "refused",
                                   education_level==9 ~ "don't know"))


sleep_df=foreign::read.xport("nhanes/P_SLQ.XPT") %>%
  janitor::clean_names() %>% 
  select(seqn,slq300,slq310,sld012,slq320,slq330,sld013,slq050,slq120) %>%
  rename(weekday_slp_time=slq300,
         weekday_wake_time=slq310,
         weekday_slp_hr=sld012,
         weekend_slp_time=slq320,
         weekend_wake_time=slq330,
         weekend_slp_hr=sld013,
         trouble_slp=slq050,
         sleepy_freq=slq120
         )              %>%
  filter(weekday_slp_time!=77777 & weekday_slp_time!= 99999) %>%
  filter(weekday_wake_time!=77777 & weekday_wake_time!= 99999) %>%
  filter(weekend_slp_time!=77777 & weekend_slp_time != 99999) %>%
  filter(weekend_wake_time!= 77777 & weekend_wake_time != 99999) %>%
  filter(trouble_slp!=7 & trouble_slp!= 9) %>%
  filter(sleepy_freq!= 7 & sleepy_freq!=9) %>%
  drop_na() %>%
  mutate(trouble_slp=case_when(trouble_slp==1 ~ "yes",
                               trouble_slp==2 ~ "no")) %>%
  mutate(sleepy_freq=case_when(sleepy_freq==0 ~ "never",
                               sleepy_freq==1 ~ "once a month",
                               sleepy_freq==2 ~ "2-4 per month",
                               sleepy_freq==3 ~ "5-15 per month",
                               sleepy_freq==4 ~ "16-30 per month",
                               )) 
  

```

```{r,echo=FALSE}
slp_df=inner_join(demo,sleep_df,by="seqn") %>%
  na_if("") %>%
  na.omit()
write_csv(slp_df, "data/slp_df.csv")
```

## National trends in short sleep duration

Short sleep duration is based on age group recommended hours of sleep per day and defined as less than 7 hours for adults. This plot shows age-adjusted prevalence of adults who reported short sleep duration from 2014 to 2020. Overall, short sleep duration prevalence was higher among males than females across years.

```{r}
year_df=read_csv("data/AdultTrends.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(cols=c("female", "male"),
               names_to = "sex",
               values_to = "age_adjust_prev") %>%
  ggplot(aes(x=year,y=age_adjust_prev,col=sex))+geom_line()+geom_point()+labs(
    title = "National trends in short sleep duration",
    x = "Year",
    y = "Age Adjusted Prevalence %"
    )

year_df+transition_reveal(year)

```
## Education Level
 
We first want to get the distribution of sleeping hours less than 7 hours across the five different education levels. We will construct a bar chart tabulating the average sleeping hours per week in each of the five education levels. Gender consideration was also added into to the bar chart in order to see a difference between female and male in each category.

```{r}
 edu_plot=slp_df %>%
  filter(weekday_slp_hr<7)%>%
  group_by(education_level,gender) %>%
  summarize(ave_sleep=mean((weekday_slp_hr*5+weekend_slp_hr*2)/7)) %>% 
  ungroup() %>%
  mutate(education_level=fct_reorder(education_level,ave_sleep)) %>%
  ggplot(aes(x=education_level,y=ave_sleep,fill=gender))+ geom_bar(width=0.5,stat="identity")+
  viridis::scale_fill_viridis(
    name = "gender",
    discrete = TRUE
  ) + geom_text(aes(label = round(ave_sleep, 2)),position = position_stack(vjust=0.9), color = "white", size = 4)+
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))+labs(
    title = "Distribution of sleeping hours across education level",
    x = "Education Level",
    y = "Average Sleeping hours per day"
    )
edu_plot
  

```

From the above plot, high school graduates has the least sleeping hours, while college graduates or above has the highest. Female and male does not have a significant difference in both groups.

## Race

A heat map was made to visualize sleeping hours less than 7 hours among different races. The below plot indicates that there is a large number of people who has a sleeping hour of 6 hours per day among the Non-Hispanic black group. There are also many Non-Hispanic Whites who has an average of 6 hours per day. 

```{r}
race_plot=slp_df %>%
  filter(weekday_slp_hr<7) %>%
  mutate(sleep_ave=(weekday_slp_hr*5+weekend_slp_hr*2)/7) %>%
  group_by(race,sleep_ave) %>%
  summarise(obs=n()) %>%
  plot_ly(
    x = ~sleep_ave, y = ~race, z = ~obs, type = "heatmap", colors = "BuPu"
  ) %>%
  colorbar(title = "Number of People", x = 1, y = 0.5) 
layout(race_plot, xaxis = list(title = "Average Sleeping Hours Per Day"), yaxis = list(title = "Race"))

```
### Race Gender Gap by Education Level

The below plot demonstr
```{r}
gender_plot=slp_df %>%
  filter(weekday_slp_hr<7) %>%
  group_by(race,education_level) %>%
  summarize(total_f=sum(gender=="female"),
            total_m=sum(gender=="male"),
            gap=total_m-total_f) %>%
  mutate(text_lable=str_c("Race=",race,"\nEducation level: ", education_level)) %>%
  plot_ly(x=~total_f,y=~total_m,text=~text_lable,color=~race,size=~gap,type="scatter",mode="markers",
          colors="viridis",sizes = c(50, 700), marker = list(opacity = 0.7))

layout(gender_plot, title = "Race Gender Gap by Education Level", xaxis = list(title = "Number of Female Sleeping less than 7 hrs"), yaxis = list(title = "Number of Male Sleeping less than 7 hrs"))

```

```{r,fig.width=10,fig.asp=.6}
income_df=slp_df %>%
  filter(weekday_slp_hr<7) %>%
  mutate(ip_stat=case_when(income_poverty_ratio > 1 ~ "not in poverty",
                           income_poverty_ratio < 1~ "in poverty",
                           income_poverty_ratio == 1~ "in poverty")) %>%
  ggplot(aes(x=weekday_slp_hr,y=ip_stat,fill=ip_stat))+
  geom_density_ridges(
    aes(point_color = ip_stat, point_shape = ip_stat,point_fill=ip_stat),
    alpha = .3, point_alpha = 0.7)+
   scale_x_continuous(
    breaks = c(2, 4, 6), 
    labels = c("2hrs", "4hrs", "6hrs"),
    limits = c(2, 6)
    )

     
  
box_plot=
  slp_df %>%
  filter(weekday_slp_hr<6) %>%
  mutate(ip_stat=case_when(income_poverty_ratio > 1 ~ "not in poverty",
                           income_poverty_ratio < 1~ "in poverty",
                           income_poverty_ratio == 1~ "in poverty")) %>%
  mutate(sleep_ave=(weekday_slp_hr*5+weekend_slp_hr*2)/7) %>%
  ggplot(aes(x=ip_stat,y=sleep_ave))+geom_boxplot(aes(fill = ip_stat), alpha = 0.3)+
  geom_hline(aes(yintercept=median(sleep_ave),
            color="red", linetype="dashed"))+
  geom_text(aes(0, median(weekday_slp_hr), label = "sleep hours median"), vjust = -0.5, hjust = 0, color = "red")

income_df+box_plot


```

```{r}
age_group= slp_df%>%
  filter(weekday_slp_hr<7) %>%
  mutate(age_gp=case_when(age>=20 & age<=30 ~ "20-30",
                          age>=31 &age <=40 ~ "31-40",
                          age>=41 &age<=50 ~ "41-50",
                          age>=51 &age<=60 ~ "51-60",
                          age>=61 &age<=70 ~ "61-70",
                          age>=71 & age <=80 ~ "71-80")) %>%
  group_by(age_gp) %>%
  summarise(ave_slp=mean((weekday_slp_hr*5+weekend_slp_hr*2)/7))%>%
  ungroup() %>%
  mutate(age_gp=fct_reorder(age_gp,ave_slp)) %>%
  ggplot(aes(x=age_gp,y=ave_slp,fill=age_gp))+ geom_bar(stat="identity")+ scale_fill_viridis_d()+
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))+
  geom_text(aes(label = round(ave_slp, 2)),position = position_stack(vjust=0.9), color = "white", size = 4)
age_group

```
